## Results


```{r}

pacman::p_load(modelsummary, gt, tidyverse, pspatreg)
```

In the previous chapter we discussed thoroughly the different areas of interest to this work. We relied on a clear framing that justifies assessing the regional diversification potential in a more contextualised and granular approach. To do that we started from the conceptual constructs of relatedness and the subsequent technology space and introduced our machine learning approach as context-rich alternative. We also introduced our coherence metric that capitalises on our approach. With that we've drawn a clear conceptual framework on how all these ideas should interact and inform on our main outcome. Essentially we propose two levels of analysis each with a clear objective but with complementary interpretation. These two analysis are opertaionalised via three models which we described their rational, specifications, and how they meet our hypothesis in the previous chapter. In this chapter we present the results of these models and contrast their interpretation to the hypothesis and the questions this work aims to address. 

In here we will decompose our finding for each of the models we specified. We will first explore the results of model 0 (@tbl-entry-ols), which treats the benchmark hypothesis and compare the empirical contribution of regional potential and relatedness to regional entry, following a large body of literature we already discussed in the literature review. Then we will present the results of model 1 (@tbl-potential-fe-ols) which explores the hypothesis on the empirical contribution of technologies' roles in the network structure from the FITS network. This model also explores the interaction between these roles and coherence to investigate whether the relationship is universal or contingent on the alignment in industrial embededdess between the region and the technology. Model 1 also extends this exploration by including the national ecosystem covariates (@tbl-potential-fe-ols-h), this to test the stability of the network metric and  their interaction as predictive factors, and also as a robustness check for their contribution compared to model 2. Essentially model 1 fits our variables to regional technological potential and assess the within-group variation. But because the national ecosystem covariates are not at a regional level, we also want to assess their contribution when we account for both within-group and between-group variation since regions are administrative artefacts that share the same national ecosystem characteristics. This nested structure motivate our model 2 presented in @tbl-potential-re. As we mentioned in the previous chapter, if the nested structure of model 2 drives most of the variation of the outcome, it could be an indication that this variation is mainly driven by spatial dependence. This is the case as we will explain later in this chapter, thus motivates our model 3. The idea here is to understand the spatial dynamics that drives regional outcomes, the test results presented in the previous chapter indicate a strong dependence on neighbouring region outcome. This justified the inclusion of the SAR model (@tbl-spatial-sar), albeit we also present the SEM model as a robustness check (@tbl-spatial-sem)



## Entry 

In @tbl-entry-ols we benchmark how much conventional relatedness density and potential (pred11) explain regional entry to a new technology. In the pooled specifications, both measures strongly predict entry: relatedness density is positive and precisely estimated ($0.055$), while Potential is larger ($0.081$). This is consistent with the interpretation that both capture diversification opportunities, however, Potential here is closer to a feasibility metric that's learned from portfolio configurations than relatedness. Once we move to the fixed-effects designs and absorb region and year fixed effects, interpretation shifts to within-region changes over time, net of common year shocks. In this setting, relatedness density remains strongly positive (e.g., $0.089$ in the density-only fixed effects model), and Potential also remains strongly positive ($0.089$ in the Potential-only fixed effects model). The coefficient of relatedness increase significantly more than potential (which remains relatively at a stable value) when we account for within-group variation in which case both relatedness and potential contribute similarly to the probability of entry. Albeit, potential seems to improve the overall and within-group fit more than relatedness (within-$R^2$ rises from $0.014$ in the density-only fixed effects model to $0.062$ in the Potential-only fixed effects model). Moreover, when we include both metrics together, the dominance of potential influence becomes clear, the coefficient of relatedness density decrease from $0.089$ to$0.028$, whereas the coefficient of potential remains stable at $0.086$. In other words, Potential appears to be relatively the dominant predictor in this horse race, while density retains residual explanatory content that may reflect a simpler proximity component that survives beyond the non-linear feasibility signal embedded in Potential. Albeit, the evidence here is not conclusive since both potential and relatedness density display a similar influence on entry when treated separately, the dominance of potential when we include relatedness density indicates that it subsumes much of the linear relatedness signal since it captures a richer context and patterns by construction compared to the linear patterns captured in the relatedness metric. 

This benchmark motivates the subsequent mechanism analysis: rather than treating relatedness as a sufficient summary statistic, we examine how portfolio alignment, network position, and accessibility structure feasibility. This move is consistent with the broader methodological critique that pairwise relatedness can miss higher-order portfolio information; for example, @tacchella2023relatedness argues that shifting from "two-product correlations" to information that captures richer configurations can improve predictive performance. This benchmark motivates our subsequent mechanism analysis: having established that Potential is relatively more informative than relatedness density in predicting entry, we now decompose Potential itself to examine how portfolio alignment (coherence), network position (betweenness, eigenvector), and national ecosystem structure feasibility.


::: {.landscape}

```{r tbl-entry-ols, echo=FALSE}
#| tbl-cap: "Regional Potential vs Relatedness density -- Entry(year + 1, LPM with fixed effects)"


table_entry_ols <- targets::tar_read(
  table_entry_ols,
  store = here::here("../../_targets_EMP")
)

table_entry_ols %>% 
  tab_options(
    table.font.size = px(9),        # font size
    data_row.padding = px(1),      # tighter body rows
    heading.padding = px(1),       # tighter title/subtitle
    column_labels.padding = px(0), # tighter header row
    table.width = pct(100)         # fill available width
  )

```

:::




## Network Effects and Coherence Moderation

As we mentioned in the beginning of this chapter, one of our main tests in our work is regarding the relationship between the technologies' network roles and regional technological potential and whether this relationship is universal or dependent on the alignment of industrial structure between regions' portfolios and a given technology. @tbl-potential-fe-ols portrait one side of the picture that we're testing here which focuses on the within-group variation.

The first observation here is that all the coefficients maintain the same sign across all specification. The coefficients of betweenness, eigenvector, coherence, and the interactions are the most notably stable in sign and significance with a slight fluctuation in size of the effect. 

As a start,  column 5  which represent our full specification shows that coherence--the alignment between a technology’s FITS embeddedness and the region’s portfolio--is the strongest predictor of potential ($0.706$ at mean betweenness and mean eigenvector)^[Because coherence interacts with both betweenness and eigenvector in the full specification, the reported coefficient is its marginal effect when both interacted variables are at their sample means. The conditional marginal effect of coherence varies with these variables but remains positive and substantive throughout the empirical distribution, as we discuss further below.]. Although regional complexity exhibits a similar predictive strength in columns 2 and 3, its effect is completely absorbed by the fixed effects.



::: {.landscape}

```{r tbl-potential-fe-ols, echo=FALSE}
#| tbl-cap: "Network Interactions -- Regional Potential (with fixed effects)"
table_pred_fe_ols_noh <- targets::tar_read(
  table_pred_fe_ols_noh,
  store = here::here("../../_targets_EMP")
)

table_pred_fe_ols_noh %>% 
   tab_options(
    table.font.size = px(8),       # font size
    data_row.padding = px(1),      # tighter body rows
    heading.padding = px(1),       # tighter title/subtitle
    column_labels.padding = px(0), # tighter header row
    table.width = pct(100)         # fill available width
  )

```

:::

Similarly, betweenness centrality, our network measure that captures bridging positions (as a proxy classification to general purpose technology role), exhibits consistent and substantive positive effects at mean coherence ($0.388$, column 5). The coherence $\times$ betweenness interaction is significantly negative ($-0.284$), but its interpretation requires evaluating the conditional marginal effect of betweenness across the coherence distribution rather than reading the interaction term in isolation^[The marginal effect of a variable on the outcome, defined as its partial derivative $\text{ME}(\tilde{x} \mid c) = \partial y / \partial \tilde{x} = \beta_{\tilde{x}} + \beta_{\tilde{x} \cdot \tilde{z}} \cdot c$, is not fixed when two z-standardised variables interact — it depends on the level $c$ at which the conditioning variable $\tilde{z}$ is evaluated. Since all variables are z-standardised, $c = 0$ is the sample mean, $c = +1$ is one standard deviation above, and $c = -1$ one standard deviation below. The effect reverses sign only if the threshold $c^* = -\beta_{\tilde{x}} / \beta_{\tilde{x} \cdot \tilde{z}}$ falls within the empirical range of $\tilde{z}$. To illustrate: the marginal effect of betweenness conditioned on coherence at one standard deviation above the mean is $\text{ME}(\text{betweenness} \mid c=+1) = 0.388 + (-0.284) \cdot (+1) = +0.104$. The same logic yields all conditional values reported in this section.]. The marginal effect of betweenness when coherence is 0 is $0.388$: at one standard deviation below mean coherence this yields $+0.672$, and at one standard deviation above mean coherence the marginal effect is $+0.104$. Bridging technologies are therefore beneficial to pursue across the full coherence distribution -- the interaction moderates the magnitude of that benefit, not its direction. The sign would only reverse at $c^* \approx +1.37$ SD, beyond the realistic empirical range. What the negative interaction establishes is a coherence-conditioned gradient: regions with weaker industrial alignment extract substantially larger returns from bridging technologies, while high-coherence regions still benefit but more modestly, consistent with their feasible opportunity set being more concentrated in already-aligned portions of the technology space^[This is further supported by the coherence $\times$ eigenvector interaction, which remains consistently non-significant. Enabling technologies are a strong and negative predictor of potential ($-0.221$, column 5), but this inhibiting effect does not vary with coherence -- the marginal effect of eigenvector is approximately $-0.221$ regardless of the region's industrial alignment. The null interaction therefore does not mean enabling technologies are irrelevant to potential; it means their structural difficulty is coherence-invariant. We will discuss this further as we progress in this chapter.] 

On the other side of this logic, the eigenvector centrality pattern presents a another story. Technologies occupying authority/prestige positions (as a proxy to key enabling technology roles) consistently inhibit potential (marginal effect $-0.221$ at mean coherence, Column 5). Its interaction with coherence (coherence $\times$ eigenvector) shows consistent non-significance across all specifications, which means the conditional marginal effect of eigenvector remains at $-0.22$ regardless of the region's industrial alignment. These technologies are therefore systematically difficult to diversify into, and coherence does not condition that difficulty. @pinheiro2025 (among others) argues in this context in favour of these results, considering that proximity alone is not a sufficient predictor of diversification. @antonietti2021going similarly argues that enabling technologies support more exploratory trajectories "only when" they are already embedded and used within the local productive system -- a condition that candidate technologies, by definition, do not yet meet at the margin. We will return to this distinction between entry difficulty and portfolio-level enabling advantage as we progress.

Our null moderation for eigenvector centrality is consistent with the idea that adopting enabling technologies is challenging and that the conditions and context for this adoption may not be fully captured by our coherence measure. For that reason we added the interaction between the network measures and the distance covariate as an robustness check. The interaction between distance and eigenvector is negative and relatively robust ($-0.080$ in columns 7 and 8). This indicates that authority technologies become particularly difficult when they are also topologically remote from any regional foothold. This mechanism is close to the account in @zhu2021strong, the authors show that diversification is more feasible through stronger, closer links. Our results contextualise this further, as this distance seems more relevant for technologies with enabling roles rather than bridging roles. This contrast can be seen in the insignificant betweenness $\times$ rawdist interaction. This (limited evidence) aligns with the intuition that general purpose technologies by construct should be accessible by construction as they bridge different areas of the network.

## National Context and Robustness

In @tbl-potential-fe-ols-h (and later on in @tbl-potential-re) we extend our framework by incorporating covariates on the national entrepreneurial ecosystem characteristics from the Global Entrepreneurship Index. The idea here is to understand whether these covariates contribute to the variation in regional potential along side the network metrics (and their interactions).

First, we can see in @tbl-potential-fe-ols-h that across the different specification the inclusion of the entrepreneurial ecosystem covariates does not change the sign, significance nor the size of the effects of the other covariates we explored in @tbl-potential-fe-ols. In fact we see a marginal improvements in the goodness of fit (within-$R^2$ rises from $0.368$ to $0.371$ and overall-$R^2$ from $0.665$ to $0.668$). Although small, this improvement shows that the inclusion of the national entrepreneurial ecosystem covariates is still contributive regardless. This stability supports a cautious conclusion that the coherence and network-position mechanisms are not simply proxies for national institutional variation. 

The contribution of these covariates, however, doesn't show the same consistency across our specification with the exception of "x13_internationalization" and "x9_competition". This is especially notable when we include the fixed effects. We already briefly discussed in the previous chapter why this could be the case here, and we will return to this point in the discussion chapter.

Regardless, Competition is consistently positive, while internationalization is consistently negative ($0.088$ and $-0.082$ respectively in Column 4). Other GEI components show heterogeneous patterns across specifications.  

This combination of stable dyadic mechanisms and non-negligible national effects is consistent with the caveat mentioned in @pinheiro2025 (also in @boschma2015towards), which cautions against treating relatedness metrics as sufficient guides to diversification arguing that institutional and policy context can shape which opportunities are realized. This heterogeneity (regardless of the consistency of competition and internationalization effects), together with the stable core mechanisms that remain consistent, constitute one of our main empirical contribution in this work. Consequently and in alignment with these observations in the literature these results provides partial evidence for H4. To explore this further model 2 provides an alternative specification that's more adequate for the hierarchical structure of the data.  


::: {.landscape}

```{r tbl-potential-fe-ols-h, echo=FALSE}
#| tbl-cap: "Network Interactions and National Ecosystem -- Regional Potential (with fixed effects)"
table_pred_fe_ols_h <- targets::tar_read(
  table_pred_fe_ols_h,
  store = here::here("../../_targets_EMP")
)

table_pred_fe_ols_h %>% 
  tab_options(
    row_group.padding.horizontal = px(0),
    table.font.size = px(9),       # font size
    data_row.padding = px(1),      # tighter body rows
    heading.padding = px(1),       # tighter title/subtitle
    column_labels.padding = px(0), # tighter header row
    table.width = pct(100)         # fill available width
  )

```

:::

## Hierarchical Structure and Spatial Implications

Similar to the extended fixed effects specification in @tbl-potential-fe-ols-h, model 2 accounts for the same covariates but considers the nested structure of the data. @tbl-potential-re employs random effects models where the regions' random intercepts are drawn from the same distribution for each country.


Crucially, the core coefficients remain stable in the hierarchical models as well. In the random effects specification comparable to the fixed effects benchmark, coherence remains positive (around $0.709$ at mean betweenness and eigenvector), and the coherence-conditioned gradient for betweenness is preserved: with a main effect of around $0.398$ at mean coherence and an interaction of $-0.292$, the marginal effect of betweenness ranges from approximately $+0.690$ at one standard deviation below mean coherence to $+0.106$ at one standard deviation above, remaining positive throughout the empirical coherence distribution. The consistency of this gradient across fixed effects and random effects specifications confirms that the coherence-conditioning of bridging benefits is not an artifact of the within-group estimation strategy but a robust structural feature of the data. Eigenvector centrality remains negative (around $-0.222$) and its interaction with coherence also remains non-significant. This further supports our previous assessment, the enabling technology inhibiting effect holds across both within-group and cross-regional specifications regardless of the coherence level. 

However and more importantly, in this model the national ecosystem covariates are uniformally stable and consistent in sign, significance and effect size. In this model we can see that variation in human capital, competition, product innovation, and high growth pillars are associated at 4%, 9%, 1.6% and 0.8% respectively with the outcome. This is in contrast to process innovation, internationalisation, and risk capital pillars whose variations are negatively associated with the outcome at 2%, 8%, and 3.5% respectively. 

These results directly address H4, and relatively confirm the intuition we formalised in this work and the results in the previous model. National entrepreneurial ecosystem characteristics matters for regional potential and explain a significant proportion of the variation (our full specification in column 5 shows the highest marginal and conditional $R^2$, 0.201 and 0.668 respectively). In fact,  this can also be observed in the high intra-class correlation ($ICC = 0.60$). This shows that $60\%$ of variance in regional potential is between-group rather than within-group (stable across the different specifications). The variance components show that country-level heterogeneity (standard deviation around $1.12$-$1.13$) exceeds region-within-country heterogeneity (approximately $1.02$–$1.03$). These decompositions inform us that, while the fixed effects estimates document robust within-region mechanisms, a substantial portion of the feasibility landscape is structured by relatively persistent differences across countries and regions.

The results of this model also help justify the spatial models that follow. The large $ICC$ and the variance decompositions indicates that regional potential is strongly structured across places. This confirms our intuition in the previous section as we introduced the rational behind model 2. The nested structure of the dyadic context matters almost as much as the dyadic components themselves. This is to say that the geography of the region (where it is, in which country) is extremely informative of its technological potential. This could justify the idea that feasibility is not independent across space but may be shaped by neighbouring structures. Consequently, the distinction between within-region fixed effects patterns and between-place random effects patterns suggests that part of the remaining heterogeneity may be geographically patterned. This is our main motivation for an explicit test for spatial dependence which is the objective of model 3 that we present its results in the next section.




::: {.landscape}

```{r tbl-potential-re, echo=FALSE}
#| tbl-cap: "Hierarchical Random Effects Model -- Regional Potential"


table_pred_re_logit <- targets::tar_read(
  table_pred_re_logit,
  store = here::here("../../_targets_EMP")
)

table_pred_re_logit %>% 
   tab_options(
    row_group.padding.horizontal = px(0),
    table.font.size = px(9),       # font size
    data_row.padding = px(0),      # tighter body rows
    heading.padding = px(1),       # tighter title/subtitle
    column_labels.padding = px(0), # tighter header row
    table.width = pct(100)         # fill available width
  )

```

::: 





## Role of space

Having established stable dyadic mechanisms at the technology–region level (Tables @tbl-potential-fe-ols and @tbl-potential-fe-ols-h) and substantial between-place heterogeneity in the hierarchical models (Table @tbl-potential-re), we test whether regional potential exhibits spatial dependence. We aggregate potential to region-year means and estimate spatial panel models with $5$-nearest-neighbor weights matrices. In alignment with the robust Lagrange Multiplier tests presented in @tbl-spatial-diagnostics in the previous chapter, we treat the SAR model as our main empirical assessment here and consider the SEM model for robustness. Essentially, this is also in alignment with the LISA map (@fig-lisa) which shows that most spatial autocorrelation is concentrated in the High-High and Low-Low quadrant of variations. 

Tables @tbl-spatial-sar and @tbl-spatial-sem reveal substantial spatial dependence. In the SAR model, the spatial autoregressive parameter $\rho$ is around $0.376$ in the interaction specification. This indicates that (average) regional potential is positively associated with neighboring regions’ outcomes. Albeit, the marginal effects of the covariates in this model should be observed in @tbl-sar-effects which describes the details of the decomposition of effect type (indirect, direct, and total effects).

In the SEM model, the spatial error parameter $\delta$ is around $0.420$ in the interaction specification, which indicates that spatial correlation could also be an unobserved components. These estimates are consistent with the interpretation that spatial context matters for regional feasibility landscapes, whether through spillovers or through correlated shocks and unobservables.

The spatial interaction terms in both models, especially in SAR, confirm the importance of space. These terms indicates that the regional feasibility landscape is spatially structured as we iterated previously when we discussed the results from the hierarchical model and the intuition behind it. However, these spatial parameters only capture correlation mechanisms and not causal structures.


```{r tbl-spatial-sar, echo=FALSE}
#| tbl-cap: "Spatial Autoregressive Model (SAR) - Regional Potential"

table_sp_sar <- targets::tar_read(
  table_sp_sar,
  store = here::here("../../_targets_EMP")
)

table_sp_sar %>% 
  tab_source_note(
    "Regional potential exhibits spatial dependence by aggregating to region-year observations and estimating spatial autoregressive (SAR) models where neighbors' potential directly affects own potential through the spatial lag parameter rho. The large and highly significant spatial lag coefficient rho = 0.38 in the interaction specification) indicates substantial positive spatial autocorrelation: regions tend to exhibit higher potential when neighboring regions also display high potential, even after controlling for regional portfolio characteristics (average coherence, betweenness, eigenvector) and complexity. In here regional aggregates describe the overall advantage of regional portfolios endowments and characteristics."
  )

```

The coefficients in @tbl-spatial-sar do not represent the marginal effects of the predictors, thus we refer to @tbl-sar-effects for this. Our focus is the interaction specification because of the conditional interpretation it provides^[In a SAR model the estimated coefficient vector $\hat{\boldsymbol{\beta}}$ does not directly recover marginal effects due to the spatial multiplier $(I - \rho W)^{-1}$. The total parametric impacts reported in @tbl-sar-effects -- derived following @lesage2009 -- decompose each covariate's total effect into a direct component (own-region response, averaging the diagonal of the partial derivative matrix $\partial \mathbf{y}/\partial \mathbf{x}_k$) and an indirect component (average spillover to all other regions, averaging the off-diagonal entries). It is these decomposed total impacts, not the raw coefficient vector, that recover the correct marginal effects and are used throughout].


The marginal effect of average portfolio betweenness centrality on regional potential is entirely conditioned by the level of average industrial coherence in the regional portfolio. From the interaction SAR specification (@tbl-sar-effects), the conditional total marginal effect of average betweenness(when average coherence = 0) is $0.0140$. This association is further decomposed into a direct own-region effect of $+0.0090$ and an indirect spillover to neighbouring regions of $+0.0050$. This picture changes substantially as coherence departs from its mean. 
At one standard deviation above mean coherence the total betweenness effect rises to $+0.0348$ (more than double the mean-coherence estimate). At one standard deviation below mean coherence the total effect of average betweenness turns negative at $-0.0068$, with both the direct ($-0.0043$) and indirect ($-0.0025$) channels reversing sign as well. This means that regions characterised with high aggregate coherence are associated with higher average potential when their portfolio's average betweenness increase by one SD, whereas for low coherence regions, higher average betweenness negatively influence their average potential. This is to say, that regions already endowed with bridging (general purpose) technologies would benefit more from higher aggregate industrial alignment on their average potential. 
The sign flip occurs at approximately $-0.67$ SD: regions below this coherence threshold experience a negative portfolio-level betweenness effect, while regions above it experience a positive one. The interaction coefficient ($+0.0208$ total) is $142\%$ of the betweenness main effect ($+0.0140$). In a non-causal sense, coherence is the governing condition that determines the direction and magnitude here. A portfolio with bridge like technologies without adequate industrial alignment is not sufficient for regions to increase their aggregate potential.To embed and activate those bridging connections into portfolio-level feasibility advantage a baseline of aggregate industrial alignment is required.




```{r fig-lisa}
#| fig-width: 8
#| fig-height: 9
#| out-width: "100%"
#| fig-dpi: 150
#| fig-cap: "Local Indicators of Spatial Association (LISA)"
#| fig-subcap: "Spatial clustering of regional potential through a LISA map based on the 2018 observations, which identifies regions where high (or low) potential clusters geographically. Regions are classified into four quadrants: High-High (HH, high potential surrounded by high-potential neighbors, typically innovation cores in Western Europe), Low-Low (LL, low potential surrounded by low-potential neighbors, typically peripheral regions in Eastern Europe or rural areas), High-Low (HL, isolated high-potential regions in low-potential neighborhoods), and Low-High (LH, lagging regions surrounded by high-potential neighbors). The map reveals that most variation concentrates in the HH and LL quadrants, consistent with the positive spatial autocorrelation documented in @tbl-spatial-sar and @tbl-spatial-sem."

lisa_map <- targets::tar_read(
  lisa_map,
  store = here::here("../../_targets_DES")
)

lisa_map +
  theme(
    plot.title = element_blank(),
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.box = "horizontal",
    legend.title = element_text(hjust = 0.5),
    legend.title.position = "top",
    legend.key.width = unit(0.1, "null"),
    legend.key.height = unit(0.5, "cm"),
    legend.box.just = "center",
    legend.title.align = 0.5
    # legend.margin = margin(t = 0, b = 0, r = 0.6, l = 0.6)
  ) 

```


The marginal effect of average portfolio eigenvector centrality complements the picture we discussed above. From the interaction SAR specification (@tbl-sar-effects), the conditional total marginal effect of average eigenvector at coherence level $\tilde{c}$ is $0.0354 + 0.0098 \cdot \tilde{coherence}$. At mean coherence ($\tilde{coherence} = 0$), a one-standard-deviation increase in average portfolio eigenvector is associated with a total effect of $+0.0354$ on average regional potential, this effect is further decomposed into a direct effect of $+0.0227$ and an indirect spillover of $+0.0127$. At one standard deviation above mean coherence, the total effect rises to $+0.0452$; at one standard deviation below, it falls to $+0.0256$. In contrast to the average betweenness coefficient, the average eigenvector effect never reverses sign across the coherence distribution. The sign-flip threshold would require coherence to be at $-3.61$ SD, well outside our current empirical range. Furthermore, the interaction coefficient ($+0.0098$ total) is $28\%$ of the eigenvector main effect ($+0.0354$ total). Thus,  coherence in the case of enabling technologies is an amplifier. This is to say that regional portfolios endowed with enabling technologies have an advantage as they are generally associated with higher potential which increases as their aggregate industrial alignment also increases. But this mechanism does not determine whether that advantage exists. This stands in direct structural contrast to betweenness, where coherence conditions the advantage of betweenness. Regions whose portfolios are concentrated in high-eigenvector technologies command higher average potential regardless of their average coherence level. But the enabling advantage of holding enabling technologies is a static feature of the regional portfolio, not a conditional one. This is also supported by the positive indirect coefficient ($+0.0127$ at mean coherence, stable at approximately $36\%$ of total across different coherence levels). This could mean that this enabling advantage is not locally contained as regions with authority-position portfolios generate feasibility spillovers to their neighbours, consistent with the argument in @antonietti2021going that enabling capacities matter when they are embedded and actively used within the local productive system rather than as isolated entry targets.


```{r message=FALSE, warning=FALSE}
#| include: false



effects_sar <- targets::tar_read(
  effects_sar,
  store = here::here("../../_targets_EMP")
)

tmp <- effects_sar %>% 
  map(summary) 

tmp$baselin %>%  names()

extract_effect_tables <- function(x) {
  purrr::imap_dfr(
    x[c("tot_table", "dir_table", "ind_table")],
    \(tab, nm) as.data.frame(tab) |>
      tibble::rownames_to_column("term") |>
      dplyr::mutate(effect = sub("_table$", "", nm), .before = 1)
  )
}

# tmp is your list of summary objects (baseline/network/interaction)
effects_tables_by_model <- purrr::map(tmp, extract_effect_tables)

full_df <- effects_tables_by_model %>% 
  purrr::imap_dfr(
    \(x,y) mutate(x, model = y)
  )

gt_tbl_simple <- full_df %>% 
  select(- "t value") %>%
  pivot_wider(names_from = "effect", values_from = c("Estimate", "Std. Error", "Pr(>|t|)"), names_sep = "__") %>% 
  select(
    model, term,
    ends_with("dir"),
    ends_with("ind"),
    ends_with("tot")
    # starts_with("Estimate__"),
    # starts_with("Std. Error__"),
    # starts_with("Pr(>|t|)__")
  ) %>% 
  gt(groupname_col = "model", rowname_col = "term") %>% 
  tab_spanner_delim(columns = 3:last_col(), delim = "__", split = "first", reverse = T) %>% 
  fmt_number(
    # columns = c("Estimate", "Std. Error", "Pr(>|t|)"),  # Specify columns to round
    decimals = 3               # Number of decimal places
  ) 


```

::: {.landscape}




```{r tbl-sar-effects, message=FALSE, warning=FALSE}
#| tbl-cap: SAR effect decomposition - Regional Potential

gt_tbl_simple %>% 
  tab_source_note(
    "This table reports the direct, indirect, and total parametric impacts of the SAR models for each specification. Because the spatial lag propagates any local shock through the neighbourhood network via the spatial multiplier, the raw coefficient vector does not recover marginal effects. The direct effect captures the own-region response, the indirect effect captures the average spillover to neighbouring regions, and the total effect is their sum."
  )
```


:::


These two conditional gradients (average betweenness and eigenvector) converge on coherence as the central organising condition of the regional feasibility landscape at the spatial level. From the interaction SAR specification (@tbl-sar-effects), average coherence carries a total effect of $+0.1044$ at mean betweenness and mean eigenvector -- the largest main effect among all portfolio-level covariates -- decomposing into a direct effect of $+0.0669$ and an indirect effect of $+0.0375$, with the indirect channel accounting for approximately $36\%$ of the total. This baseline effect is itself modulated by both betweenness and eigenvector through their respective interactions, but what the conditional gradients in paragraphs above make precise is the asymmetry in how coherence operates across the two network dimensions. First, for betweenness, coherence is the condition that determines whether bridging technologies is an asset or a liability at the portfolio level. Second, for eigenvector, coherence amplifies an advantage that exists independently of alignment. In both cases, however, the direction of the coherence effect on regional potential is unambiguous and stable -- higher industrial alignment is associated with higher average potential across the full distribution of betweenness and eigenvector values, with the indirect effects confirming that this alignment advantage propagates spatially to neighbouring regions. Coherence is therefore not only a dyadic feasibility correlate, as the earlier models established, but the structural condition that governs how portfolio-level network positions -- bridging and enabling alike -- translate into regional potential and its spatial diffusion.


```{r tbl-spatial-sem, echo=FALSE}
#| tbl-cap: "Spatial Error Model (SEM) - Regional Potential"


table_sp_sem <- targets::tar_read(
  table_sp_sem,
  store = here::here("../../_targets_EMP")
)

table_sp_sem %>% 
  tab_source_note(
    source_note = "This table estimates a spatial error model (SEM) as a robustness check to distinguish between substantive spatial interdependence (neighbors' outcomes affecting own outcome, as in SAR) and spatially correlated unobservables (common shocks or omitted factors that cluster geographically). The spatial error parameter lambda is positive and significant (0.42 in the interaction specification). Aafter controlling for observable regional characteristics, residuals remain spatially autocorrelated. However, the robust Lagrange Multiplier diagnostics reported in the previous chapter favor the SAR specification. The regional covariate patterns in SEM closely mirror those in SAR."
    )
  

```

## Summary 


In this chapter we presented the results of our empirical analysis which we introduced in the first section. The second section addressed our benchmark question of whether regional technological potential is a more informative predictor of entry than relatedness density. Our results show that potential subsumes much of the relatedness density signal and provides marginal improvements in model fit. Although potential does not clearly dominate relatedness density when both are included as co-predictors, the evidence is consistent with our conservative criterion -- potential contributes more across specifications and there is no specification in which relatedness density dominates potential. We therefore accept H1, while acknowledging that a more conclusive comparison would be beneficial.

The third section presented the results of model 1 concerning the influence of technology characteristics and portfolio alignment on regional technological potential. Coherence is consistently the strongest predictor of potential across all specifications, confirming H3a -- regions with more coherent portfolios face lower feasibility barriers across the technology space. Regarding technology characteristics, enabling technologies are consistently associated with lower potential regardless of the level of portfolio alignment, confirming H2b. This inhibiting effect is stable and does not vary with coherence, meaning H3c is not supported -- the structural difficulty of entering foundational technologies appears to operate through mechanisms that portfolio alignment alone does not address. For bridging technologies, the picture is more nuanced. They are consistently associated with higher potential, confirming H2a, and this association is stronger for regions with lower coherence, confirming H3b. The advantage of bridging technologies diminishes as coherence increases but remains positive throughout the empirical distribution, meaning they are a reliable diversification pathway across the coherence spectrum, with the largest returns for regions where direct alignment is limited.

The forth section introduced national ecosystem covariates. In the fixed effects model the results were not fully consistent across specifications, which is expected given the identification constraints of that design. The hierarchical model addressed this directly and showed clear and consistent effects across all ecosystem dimensions, confirming H4. National institutional and market characteristics contribute independently to regional potential beyond what portfolio mechanisms explain, though the direction of effects varies across ecosystem dimensions.

The fifth section presented the spatial model results. The spatial lag parameter is positive and significant, confirming H5a -- regional feasibility landscapes are spatially interdependent rather than locally contained. At the aggregate portfolio level, the role of coherence, bridging, and enabling characteristics shifts relative to the dyadic results. For portfolios concentrated in enabling technologies, the association with average potential is positive and amplified by coherence, consistent with H5c. This contrasts with the dyadic finding where enabling technologies inhibit potential -- once they exist in the portfolio, their structural position in the capability network supports the feasibility of other technologies by construction, which is reflected in higher aggregate potential. For portfolios concentrated in bridging technologies, the association with average potential is conditional on coherence, partially confirming H5b. Below the coherence threshold the association is negative, and above it the association is positive. This means that bridging technologies as targeted diversification options are reliable at the dyadic level, but their contribution to aggregate portfolio potential requires a baseline of industrial alignment to materialise. Bridging characteristics alone are therefore not a sufficient portfolio-level instrument -- their aggregate value depends on the broader industrial structure in which they are embedded.




<!--




```{r tbl-potential-fe-logit, eval=FALSE}
#| tbl-cap: "logit(pred1) - logistic"
table_pred_fe_logistic_noh <- targets::tar_read(
  table_pred_fe_logistic_noh,
  store = here::here("../../_targets_EMP")
)

table_pred_fe_logistic_noh

```




```{r tbl-potential-fe-logit-h, eval=FALSE}
#| tbl-cap: "logit(pred1) - logitic"
table_pred_fe_logistic_h <- targets::tar_read(
  table_pred_fe_logistic_h,
  store = here::here("../../_targets_EMP")
)

table_pred_fe_logistic_h

```







```{r tbl-potential-re1, eval=FALSE}
#| tbl-cap: "pred1 - ols"
table_pred_re_ols <- targets::tar_read(
  table_pred_re_ols,
  store = here::here("../../_targets_EMP")
)

table_pred_re_ols

```


-->



